---
title: "Data Science 1 Homework 1"
author: "Samina Rashiq"
date: "2024-09-21"
output: github_document
---
```{r tidyverse, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE}
library(tidyverse)
```
# Problem 1

#### Load penguins dataset
```{r penguins}
data("penguins", package = "palmerpenguins")
```

```{r echo=FALSE}
# Name important variables
num_rows <- nrow(penguins)
num_cols <- ncol(penguins)

# Calculate mean flipper length
mean_flipper_length <- mean(penguins$flipper_length_mm, na.rm = TRUE)
```

#### Describe penguins dataset
```markdown
The`penguins` dataset contains information about many types of penguins including the variables:
- `species`: species of penguin (Adelie, Chinstrap,or Gentoo)
- `island`: island where penguing was observed (Biscoe, Dream, or Torgersen)
- `bill_length_mm` : length of penguin's bill in mm (mean = `r mean(penguins$bill_length_mm, na.rm = TRUE)`)
- `bill_depth_mm`: depth of penguin's bill in mm (mean = `r mean(penguins$bill_depth_mm, na.rm = TRUE)`)
- `flipper_length_mm`:length of penguin's flipper in mm (mean = `r mean(penguins$bill_depth_mm, na.rm = TRUE)`). 
- `body_mass_g`: body mass of penguin in g (mean = `r mean(penguins$body_mass_g, na.rm = TRUE)`)
- `sex`: The sex of the penguin in (male or female)
- `year`: The year data was collected.

The `penguins` dataset contains `r num_rows` rows and `r num_cols` columns. The mean flipper length is `r mean_flipper_length` mm. 

```
#### Make a scatterplot of flipper length vs. bill length
```{r yx_scatter, echo = FALSE, warning=FALSE, message=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(
    title = "Flipper length vs. bill length",
    x = "Bill length (mm)",
    y = "Flipper length (mm)",
  )

#Saving plot as image
ggsave(
  "scatterplot1.png",
  plot = last_plot ()
)

```

# Problem 2

#### Create data frame
```{r df_create}
set.seed(420)

problem2_df = tibble(
  norm_samp = rnorm(10, mean = 0, sd = 1),
  vec_logical = norm_samp > 0,
  vec_char = sample(state.name, 10, replace = TRUE),
  vec_factor = factor(
    sample(
      c("lvl1", "lvl2", "lvl3"), 10, replace = TRUE))
)

print(problem2_df)
```

#### Calculate the mean of each variable
```{r}
mean(pull(problem2_df, norm_samp))

mean(pull(problem2_df, vec_logical))  

mean(pull(problem2_df, vec_char)) 

mean(pull(problem2_df, vec_factor))
```

#### Convert logical, character, and factor variables to numeric variables

```{r means_calc,}
num_logical <- as.numeric(pull(problem2_df, vec_logical))

num_char <- as.numeric(pull(problem2_df, vec_char))

num_factor <- as.numeric(pull(problem2_df, vec_factor))

```
R treats logical values as TRUE = 1 and FALSE = 0, which is why a mean could be calculated before conversion. 

R can't change character variables into numeric values because they are not binary or ordinal. This is why a mean could not be calculated and conversion results in error. 

Factor variable could not have mean calculated initially because integer codes were not assigned. The conversion did that, allowing the calculation to occur.









